\vspace{-2.0cm} \textbf{Talking with signs} \\ \Large \textbf{A simple method to detect nouns and numbers in a non-annotated signs language corpus}

### 2019-12-03 article/article-tesis.tex

\title{%
	\vspace{-2.0cm}
	\textbf{Talking with signs} \\	
	\Large \textbf{A simple method to detect nouns and numbers in a non annotated signs language corpus}
}

@@

\begin{abstract}
People with deafness or hearing disabilities who aim to use computer based systems relies on state of art video classification and human action recognition techniques that combine traditional movement pattern recognition and deep learning techniques. In this work we present a pipeline for semi automatic video annotation applied to a non annotated \textit{Peruvian Signs Language} (PSL) corpus along with a novel method for a progressive detection of PSL elements (nSDm). We produced a set of video annotations indicating signs appearances for a small set of nouns and numbers along with a \textit{labeled PSL dataset} (PSL dataset). A composite model obtained from the combination of a 2D CNN trained with movement patterns extracted from the PSL dataset using Lucas Kanade optical flow, and a RNN with LSTM cells trained with raw rgb frames extracted from the PSL dataset achieved 0.5994 AUC for the Presission-Recall curve on sings classification tasks, reporting state of art results over the PSL dataset. \\
\keywords{Video Classification, Human Actions Detection, Peruvian Signs Language, Optical Flow, 2D CNN, LSTM.}
\end{abstract}

@@

The Peruvian Institute of Informatics and Statistics (INEI) conducted a national disabilities survey with the objective of segmenting and acquire a better understanding about disabilities that affect the Peruvian population \cite{disabilities_survey_2012}, the results showed that 1.8\% of the Peruvian population suffer at least partial when not permanent deafness or hearing limitations.

Peruvians with deafness or hearing limitations use the Peruvian Signs Language (PSL) as their main communication medium. PSL is of mandatory usage at universities and certain public institutions, here the importance of designing systems that are capable to support PSL inputs and outputs. Furthermore, in the same way as spoken languages, signs languages also present local variations e.g. people who live in Lima metropolitan area are not expected to use the same set of signs as people in other parts of the territory. This work uses the PSL variation used in Lima due to the difficulty or inability to find datasets for other PSL variations.

\textit{The Grammar and Signs research group of the Pontifical Catholic University of Peru (PUCP)} built the first PSL corpus \cite{lsp_dataset} which is publicly available at the university digital archives, it is important to highlight that the corpus is neither labeled or annotated and cannot be used as it is for training or testing a model.

In this work we are approaching signs detection as a supervised learning task. Supervised learning requires labeled dataset to achieve satisfactory results during training and inference tasks. At the time of writing this work there were no labeled datasets available for PSL \cite{lsp_2015}, it configures a gap that could prevent or hinder research work on Human-Computer-Interaction at the Peruvian or Latin American space.

@@

\begin{enumerate}[(i)]
\item What are the currently available techniques for producing a labeled dataset for a set of signs limited to nouns and number from the non annotated PSL corpus?\label{q1}
\item What are most relevant and currently available techniques for training a model with the labeled dataset described in the question above for detecting PSL nouns and numbers?\label{q2}
\item How precise and exhaustive is the model described in the above question on the detection of PSL nouns and numbers?\label{q3}
\end{enumerate}

This work has the main objective of producing a simple method that can be used as a baseline for other researchers interested on studying signs language and their different applications on the Human-Computer-Interaction field, we believe this work will produce a positive impact on the artificial intelligence community towards to an increase in the number of research works using PSL which can contribute to increasing the number of people with deafness or hearing disabilities that can use computer based systems. We have divided this work main objective into following specific objectives for better traceability:

\begin{enumerate}[(i)]
\item Produce a labeled PSL dataset limited to nouns and number\label{o1}
\item Design a train a novel signs detection model (nSDm) for detecting signs at the labeled PSL dataset\label{o2}
\item Determine what is the performance of nSDm in terms of precision and recall\label{o3}
\end{enumerate}

The rest of the article is organized as follows. In section \ref{relatedwork} we review the related work on video classification for human actions recognition using network architectures that combine CNNs, 3D CNNs and movement patterns  for better features learning, we also review state of art pose estimation techniques. In section \ref{method} we introduce nSDm describing its design and architecture. In section \ref{experimentation} we evaluate nSDm precision and recall and answer research questions \ref{q1},\ref{q2},\ref{q3}. In section \ref{datasetdesc} we describe the PSL dataset produced at PUCP. In section \ref{videoannot} we describe the video annotation and data pre-processing techniques applied to produce the labeled PSL dataset and finally in section \ref{conclusion} we present our conclusions and future work.
\section{Related Work} \label{relatedwork}
\subsection{Action Recognition}
Human action recognition is a extensively studied field, action recognition dataset like UCF101, HMDB51, THUMOS14  are available, researches tried to solve the human action recognition problem using different approaches including Optical Flow and 3D CNN \cite{qiu2017learning}.  \textbf{Optical Flow,} is defined as the pattern obtained from the motion of objects, surfaces and edges in a visual scene caused by the relative motion between the observer and a scene, it is achieved by distributing movement velocities and brightness across frames. It is a  key concept in action recognition from videos \cite{wang2019hallucinating}. Optical flow estimation is treated as an image reconstruction problem. Given a frame set, the optical flow is generated and allows to reconstruct one frame from the others \cite{zhu2018hidden}. Formally, taking the optical flow displacement field as input and training a CNN with it, then the network should have learned useful representations of the underlying motions. Even though Optical Flow represents the movement between a set of frames, if camera motion is considered as an action motion, it may corrupt the action classification \cite{wang2013dense}. Various types of camera motion can be observed in realistic videos, e.g., zooming, tilting, rotation, etc. \textbf{Motion Boundary Histogram (MBH),} It is a simple an efficient way to achieve robustness during human action detection when camera movements are mixed within the recorded actions by computing derivatives separately for the horizontal and vertical components of the optical flow. Since MBH represents the gradient of optical flow, locally constant camera motion is removed and information about changes in the flow field is kept. MBH is more robust to camera motion than optical flow, thus more discriminative for action recognition.\cite{wang2013dense}. 3D CNN are not as effective as optical flow to detect human actions on its own, 3D CNN can be trained to learn optical flow so we can avoid costly computation and storage and obtain task-specific motion representation  \cite{zhu2018hidden} and increase models performance, precision and exhaustiveness on human action recognition.
\subsection{Pose Estimation}
Pose estimation is also an extensively studied field. Techniques based on key points have shown state of art results on human pose estimation. An approach on key points estimation \cite{tulsianiMalik} uses \textit{Point of View Determination} and \textit{Key Points Prediction} components. Point of View Determination is formulated by the prediction of three Euler angles (azimut, elevation and cyclotation) generating a global position estimate, then a local appearance is modeled by obtaining a heat map that corresponds to the verisimilitude spatial distribution for each key point, finally key points predictions are obtained by combining heat maps obtained in a previous stage with a conditioned verisimilitude at the point of view predicted in the previous stage.

Key points detection methods based CNNs have received an special attention in Human Pose Detection problems, CNNs methods are divided in bottom-up and top-down. Bottom-up methods process images from low resolution  to high resolution, focusing first on detecting joints and then before associating them to human actions. Top-down methods process focused first on detecting humans and then estimating the human pose to predict key points.

The datasets MPII and COCO have been used in state of art methods obtaining good results\cite{XiaoWuWeiSimpleBaseline} and establishing a framework for future work in combination with classic approaches like optical flow for recognizing patterns movement between frames by increasing accuracy on key points detection.
\subsection{Video Classification}
\textbf{Bag of Words (BoW)} or \textbf{Bag of Visual Words (BoVW)} based on natural language processing techniques is one of the simplest and oldest local descriptor encoding strategies. In its simplest form, it consists of (i) clustering with k-means a collection of descriptor vectors from the training set to build so-called visual vocabulary, (ii) as signing each descriptor to its nearest cluster center from the visual dictionary, and (iii) aggregating the one-hot assignment vectors via average pooling \cite{wang2019hallucinating}, when applied to Computer Vision is a technique used to create images representations or features vectors used that can be learned by CNNs, resulting on improved images classification and video classification. 
Feature trajectory detection are much improved using statistical methods like \textbf{Fisher Vectors} obtaining better results over traditional BoW \textbf{Fussing parallel CNN.}, the Bag of Visual Words representation suffers from sparsity and high dimensionality, in the other hand representations obtained using the Fisher Vectors kernel are more compact and dense which results on better results on image and video classification problems.
\section{Method}\label{method}
\subsection{Video Annotation}\label{videoannot}
The PSL dataset is non annotated because there is not a direct relation between the instant when a sign is emitted and when its translation to Spanish is delivered, we propose a semi automatic video annotation pipeline described in Figure \ref{fig:video-annotation-pipeline} for cleaning, pre-processing and analyzing PSL videos in order to produce an labeled PSL dataset that can be used for training nSDm using supervised learning. The pipeline is described in detail in sections  \ref{manual-video-cleanup}, \ref{video-pre-processing}, \ref{audio-transcription-analysis} and \ref{samples-generation}

@@

\subsubsection{Manual Automatic Video Clean Up}\label{manual-video-cleanup}
The PSL recordings described on \ref{datasetdesc} contain a considerable amount of noise introduced during recording sessions. It makes difficult to easily find video intervals that clearly show a relation between signs emitted by the informant and the translation delivered by the translator, the noise factors are listed in the following list:
\begin{itemize}
	\item Multiple participants speaking during the session.
	\item Conversations between participants that are not relevant to emitted sings.
	\item High frequency of large silent periods.
\end{itemize}
A manual video cleanup process is required to find noise free video intervals, it is a process that requires watching all the videos available at the PSL corpus for manually annotating the instant when an informant started emitting sings along with the instant when the translator delivered a translation, table \ref{table:noise-free-video-segments} shows a manual annotation example.

@@

\subsubsection{Video Pre-Processing}\label{video-pre-processing}
Non annotated PSL videos require processing before any metadata can be extracted, we propose a sequence of pre-processing tasks that take advantage of the annotation generated on \ref{manual-video-cleanup}. A \textbf{video splitting processor} generates a set of video chunks using the \textit{ffmpeg} multimedia framework and stores produced video chunks in Amazon S3 for later usage. Audio within video chunks is then transcribed by an \textbf{audio transcription processor}, using the Amazon Transcription service, we selected the Amazon Transcription service because it provides an accurate mapping between audio participants and transcribed words along with useful metadata that describes the start and end time when words are pronounced by the translator.

At the moment of writing this work Amazon Transcription service only supported Spain and US Spanish, it caused certain words that are specific for Peruvian Spanish not being fully recognized, in order to improve transcription accuracy we built a custom vocabulary containing Peruvian expressions which improved  Peruvian words recognition, for the matters of this work Peruvian words that remained unrecognized were omitted and not processed.

@@

A \textbf{Optical Flow Samples generation processor} uses video frames and the hierarchical folder structure generated by the RGB samples generation processor to calculate an Optical Flow representation for PSL elements and store them in a hierarchical folder structure that will also used by the nSDm data loaders to feed the optical flow branch on the nSDm model architecture \ref{nsdm-architecture} during training and testing. We selected optical flow as a samples generation strategy because its ability to represent movement traces from previous frames, it is particular useful for representing body movement patterns executed by informant while emitting a PSL sign, a PSL sign is made up of different body movements including: elbow, arms, neck, eyes, shoulders and hands, which are performed quickly, a way to detect movement traces between frames allows to generate a single image representation of all movement involved on a sign emitting (Figure \ref{fig:opticalflow-two}).

@@

Implementation details can be found at https://github.com/erichuizapucp/signs-recognition
\subsubsection{Architecture}\label{nsdm-architecture}
A series of models stacked in a composed architecture focused on learning visual and time series features using two network architecture flavors enforced with a flexible data input pipeline for data feeding as shown in Figure \ref{fig:ndsm-architecture}, transformation and normalization. The \textbf{Input pipeline} accesses the labeled PSL samples and applies transformations preparing the data for upper layers, transformations were applied to both optical flow and RGB frames, PSL samples were resized to be compliant with ImageNet pre-trained models using a 224 by 224 shape and three channels for color images, data augmentation transformation were not applied due to the nature of the experiment where samples were captured using similar light conditions and camera orientations, PSL samples were transformed to tensors and normalized to floats in the [0, 1] interval, finally the transformed and normalized versions of optical flow and RGB samples were tight together in a tensor structure along with the label for usage at upper layers.
\begin{table}[!htb]
\captionsetup{size=footnotesize}
\begin{tabular}{ p{16em} c}
\toprule
\multicolumn{1}{c}{\textbf{Hyper parameter}} & 
	\multicolumn{1}{c}{\textbf{Value}}\\
\midrule
\textbf{Learning Rate}&	0.001\\
\textbf{No Epochs}&	5\\
\textbf{No Steps per Epoch}&	5\\
\textbf{Batch Size}&	10\\
\textbf{Shuffle Buffer Size}&	50\\
\bottomrule
\end{tabular}
\caption{nSDm training hyper parameters} \label{tab:nSDm-hyperparams}
\end{table}
The \textbf{Wrapper Layer} wraps a CNN network designed for learning visual features from optical flow and a RNN network designed for learning features from series of video frames, the layer receives data from the input pipeline then splits it into optical flow and RGB samples and passes them to the CNN and RNN networks respectively, the logits  produced are then concatenated and passed to the dense layer.
The dense layer uses a softmax activation function to convert logits into probabilities used for a correct sign classification. Hyper parameters used during the training process are listed on Table \ref{tab:nSDm-hyperparams}

@@

The \textbf{Opticalflow model} described in Figure \ref{fig:opticalflow-architecture} receives decoded opticalflow samples, it uses a Resnet152 backbone pre-trained with ImageNet, we used a fine tuning transfer learning approach, the backbone produces a 7x7x2048 output that then is passed to a Global Average Pooling layer for obtaining a flattened output of 1x1x2048 which is then passed to a dense layer for only logits generation.

@@

The \textbf{Recurrent RGB model} described in Figure \ref{fig:rgb-architecture} receives a sequence of decoded video frames, it uses a \textit{many to one} architecture with LSTM cells that hold state of 2048 length, the output produced by the recurrent layers are then passed to a dense layer for only logits generation.

@@

\subsection{Sign detection testing}\label{new-videos-testing}
We selected videos for consultant 24 \ref{datasetdesc} for testing nSDm (the model was not trained with consultant 24 videos), we proceed to extract RGB frames for video fragments of five seconds length moving ahead one second on each loop until the end of the video that way we can reduce the number of signs cut or incomplete between video fragments. Optical flow and RGB frames generators feed nSDm for detecting signs \ref{nsdm} present in the video fragment.

@@

\section{Experimentation}\label{experimentation}
\subsection{Dataset Description \cite{lsp_dataset}}\label{datasetdesc}
The PSL dataset was developed by the PUCP Grammar and Signs research group in 2014 and consists in a set of videos recorded during the interviews of 24 individuals, 12 male and 12 female informants, all of them are Lima Peru residents and reported to be borne with a permanent deafness condition or acquired the condition before the acquisition of Spanish.

>>>>

\title{%
	\vspace{-2.0cm}
	\textbf{Talking with signs} \\	
	\Large \textbf{A simple method to detect nouns and numbers in a non-annotated signs language corpus}
}

@@

\begin{abstract}
People with deafness or hearing disabilities who aim to use computer based systems rely on state-of-art video classification and human action recognition techniques that combine traditional movement pattern recognition and deep learning techniques. In this work we present a pipeline for semi-automatic video annotation applied to a non-annotated \textit{Peruvian Signs Language} (PSL) corpus along with a novel method for a progressive detection of PSL elements (nSDm). We produced a set of video annotations indicating signs appearances for a small set of nouns and numbers along with a \textit{labeled PSL dataset} (PSL dataset). A composite model obtained from the combination of a 2D CNN trained with movement patterns extracted from the PSL dataset using Lucas Kanade optical flow, and a RNN with LSTM cells trained with raw rgb frames extracted from the PSL dataset achieved 0.5994 AUC for the Precision-Recall curve on signs classification tasks, reporting state-of-art results over the PSL dataset. \\
\keywords{Video Classification, Human Actions Detection, Peruvian Signs Language, Optical Flow, 2D CNN, LSTM.}
\end{abstract}

@@

The Peruvian Institute of Informatics and Statistics (INEI) conducted a national disabilities survey with the objective of segmenting and acquiring a better understanding about disabilities that affect the Peruvian population \cite{disabilities_survey_2012}. Results showed that 1.8\% of the Peruvian population suffer at least partial when not permanent deafness or hearing limitations.

Peruvians with deafness or hearing limitations use the Peruvian Signs Language (PSL) as their main communication medium. PSL is of mandatory usage at universities and certain public institutions, henceforth the importance of designing systems that are capable to support PSL inputs and outputs. Furthermore, in the same way as spoken languages, signs languages also present local variations e.g. people who live in Lima metropolitan area are not expected to use the same set of signs as people in other parts of the territory. This work uses the PSL variation used in Lima due to the difficulty or inability to find datasets for other PSL variations.

\textit{The Grammar and Signs research group of the Pontifical Catholic University of Peru (PUCP)} built the first PSL corpus \cite{lsp_dataset} which is publicly available at the university digital archives. It is important to highlight that the corpus is neither labeled or annotated and cannot be used as it is for training or testing a model.

In this work we are approaching signs detection as a supervised learning task. Supervised learning requires labeled datasets to achieve satisfactory results during training and inference tasks. At the time of writing this work there were no labeled datasets available for PSL \cite{lsp_2015}. It configures a gap that could prevent or hinder research work on Human-Computer-Interaction at the Peruvian or Latin American space.

@@

\begin{enumerate}[(i)]
\item What are the currently available techniques for producing a labeled dataset for a set of signs limited to nouns and number from the non-annotated PSL corpus?\label{q1}
\item What are most relevant and currently available techniques for training a model with the labeled dataset described in the question above for detecting PSL nouns and numbers?\label{q2}
\item How precise and exhaustive is the model described in the above question on the detection of PSL nouns and numbers?\label{q3}
\end{enumerate}

This work has the main objective of producing a simple method that can be used as a baseline for other researchers interested on studying signs language and their different applications on the Human-Computer-Interaction field, we believe this work will produce a positive impact on the artificial intelligence community towards an increase in the number of research works using PSL which can contribute to increasing the number of people with deafness or hearing disabilities that can use computer based systems. We have divided this work main objective into following specific objectives for better traceability:

\begin{enumerate}[(i)]
\item Produce a labeled PSL dataset limited to nouns and numbers\label{o1}
\item Design and train a novel signs detection model (nSDm) for detecting signs at the labeled PSL dataset\label{o2}
\item Determine what is the performance of nSDm in terms of precision and recall\label{o3}
\end{enumerate}

The rest of the article is organized as follows. In section \ref{relatedwork} we review the related work on video classification for human actions recognition using network architectures that combine CNNs, 3D CNNs and movement patterns  for better features learning, we also review state-of-art pose estimation techniques. In section \ref{method} we introduce nSDm describing its design and architecture. In section \ref{experimentation} we evaluate nSDm precision and recall and answer research questions \ref{q1},\ref{q2},\ref{q3}. In section \ref{datasetdesc} we describe the PSL dataset produced at PUCP. In section \ref{videoannot} we describe the video annotation and data pre-processing techniques applied to produce the labeled PSL dataset and finally in section \ref{conclusion} we present our conclusions and future work.
\section{Related Work} \label{relatedwork}
\subsection{Action Recognition}
Human action recognition is an extensively studied field. Action recognition dataset like UCF101, HMDB51, THUMOS14  are available, researches tried to solve the human action recognition problem using different approaches including Optical Flow and 3D CNN \cite{qiu2017learning}.

\textbf{Optical Flow,} is defined as the pattern obtained from the motion of objects, surfaces and edges in a visual scene caused by the relative motion between the observer and a scene. It is computed by distributing movement velocities and brightness across frames. It is a  key concept in action recognition from videos \cite{wang2019hallucinating}. Optical flow estimation is treated as an image reconstruction problem. Given a frame set, the optical flow is generated and allows to reconstruct one frame from the others \cite{zhu2018hidden}. Formally, taking the optical flow displacement field as input and training a CNN with it, then the network should have learned useful representations of the underlying motions. Even though Optical Flow represents the movement between a set of frames, if camera motion is considered as an action motion, it may corrupt the action classification \cite{wang2013dense}. Various types of camera motion can be observed in realistic videos, e.g., zooming, tilting, rotation, etc.

\textbf{Motion Boundary Histogram (MBH)} is a simple an efficient way to achieve robustness during human action detection when camera movements are mixed within the recorded actions by computing derivatives separately for the horizontal and vertical components of the optical flow. Since MBH represents the gradient of optical flow, locally constant camera motion is removed and information about changes in the flow field is kept. MBH is more robust to camera motion than optical flow, thus more discriminative for action recognition.\cite{wang2013dense}. 3D CNN are not as effective as optical flow to detect human actions on its own, 3D CNN can be trained to learn optical flow so we can avoid costly computation and storage and obtain task-specific motion representation  \cite{zhu2018hidden} and increase models performance, precision and recall on human action recognition.
\subsection{Pose Estimation}
Pose estimation is also an extensively studied field. Techniques based on key points have shown state-of-art results on human pose estimation. An approach on key points estimation \cite{tulsianiMalik} uses \textit{Point of View Determination} and \textit{Key Points Prediction} components. Point of View Determination is formulated by the prediction of three Euler angles (azimut, elevation and cyclotation) generating a global position estimate, then a local appearance is modeled by obtaining a heat map that corresponds to the spatial distribution likelihood for each key point, finally key points predictions are obtained by combining heat maps obtained in a previous stage with a conditioned likelihood at the point of view predicted in the previous stage.

Key points detection methods based CNNs have received an special attention in Human Pose Detection problems. CNNs methods are divided in bottom-up and top-down. Bottom-up methods process images from low resolution  to high resolution, focusing first on detecting joints before associating them to human actions. Top-down methods focus first on detecting human subjects and then estimating the human pose to predict key points.

The datasets MPII and COCO have been used in state-of-art methods obtaining good results\cite{XiaoWuWeiSimpleBaseline} and establishing a framework for future work in combination with classic approaches like optical flow for recognizing patterns movement between frames by increasing accuracy on key points detection.
\subsection{Video Classification}
\textbf{Bag of Words (BoW)} or \textbf{Bag of Visual Words (BoVW)} based on natural language processing techniques is one of the simplest and oldest local descriptor encoding strategies. In its simplest form, it consists of (i) clustering with k-means a collection of descriptor vectors from the training set to build so-called visual vocabulary, (ii) as signing each descriptor to its nearest cluster center from the visual dictionary, and (iii) aggregating the one-hot assignment vectors via average pooling \cite{wang2019hallucinating}, when applied to Computer Vision is a technique used to create images representations or features vectors used that can be learned by CNNs, resulting on improved images classification and video classification. 
Feature trajectory detection are much improved using statistical methods like \textbf{Fisher Vectors} obtaining better results over traditional BoW \textbf{Fussing parallel CNN.}. The Bag of Visual Words representation suffers from sparsity and high dimensionality, in the other hand representations obtained using the Fisher Vectors kernel are more compact and dense which results on better results for image and video classification problems.
\section{Method}\label{method}
\subsection{Video Annotation}\label{videoannot}
The PSL dataset is non-annotated because there is not a direct relation between the instant when a sign is emitted and when its translation to Spanish is delivered. We propose a semi-automatic video annotation pipeline described in Figure \ref{fig:video-annotation-pipeline} for cleaning, pre-processing and analyzing PSL videos in order to produce an labeled PSL dataset that can be used for training nSDm using supervised learning. The pipeline is described in detail in sections  \ref{manual-video-cleanup}, \ref{video-pre-processing}, \ref{audio-transcription-analysis} and \ref{samples-generation}

@@

\subsubsection{Manual Automatic Video Clean Up}\label{manual-video-cleanup}
The PSL recordings described on \ref{datasetdesc} contain a considerable amount of noise introduced during recording sessions. It makes difficult to easily find video intervals that clearly show a relation between signs emitted by the informant and the translation delivered by the translator. Noise factors are the following:
\begin{itemize}
	\item Multiple participants speaking during the session.
	\item Conversations between participants that are not relevant to emitted sings.
	\item High frequency of large silent periods.
\end{itemize}
A manual video cleanup process is required to find noise free video intervals. This process requires watching all the videos available at the PSL corpus for manually annotating the instant when an informant started emitting sings along with the instant when the translator delivered a translation, Table \ref{table:noise-free-video-segments} shows a manual annotation example.

@@

\subsubsection{Video Pre-Processing}\label{video-pre-processing}
Non-annotated PSL videos require processing before any metadata can be extracted, we propose a sequence of pre-processing tasks that take advantage of the annotation generated on \ref{manual-video-cleanup}. A \textbf{video splitting processor} generates a set of video chunks using the \textit{ffmpeg} multimedia framework and stores produced video chunks in Amazon S3 for later usage. Audio within video chunks is then transcribed by an \textbf{audio transcription processor}, using the Amazon Transcription service, we selected the Amazon Transcription service because it provides an accurate mapping between audio participants and transcribed words along with useful metadata that describes the start and end time when words are pronounced by the translator.

At the moment of writing this work Amazon Transcription service only supported Spain and US Spanish. This caused certain words that are specific for Peruvian Spanish not being fully recognized, in order to improve transcription accuracy we built a custom vocabulary containing Peruvian expressions which improved  Peruvian words recognition, for the matters of this work Peruvian words that remained unrecognized were omitted and not processed.

@@

An \textbf{Optical Flow Samples generation processor} uses video frames and the hierarchical folder structure generated by the RGB samples generation processor to calculate an Optical Flow representation for PSL elements and store them in a hierarchical folder structure that will also be used by the nSDm data loaders to feed the optical flow branch on the nSDm model architecture \ref{nsdm-architecture} during training and testing. We selected optical flow as a samples generation strategy due to its ability to represent movement traces from previous frames. It is particular useful for representing body movement patterns executed by informant while emitting a PSL sign. A PSL sign is made up of different body movements including: elbow, arms, neck, eyes, shoulders and hands, which are performed quickly, a way to detect movement traces between frames allows to generate a single image representation of all movement involved on a sign emission(Figure \ref{fig:opticalflow-two}).

@@

Implementation details can be found at https://github.com/erichuizapucp/signs-recognition
\subsubsection{Architecture}\label{nsdm-architecture}
A series of models stacked in a composed architecture focused on learning visual and time series features using two network architecture flavors enforced with a flexible data input pipeline for data feeding as shown in Figure \ref{fig:ndsm-architecture}, transformation and normalization.

The \textbf{Input pipeline} accesses the labeled PSL samples and applies transformations preparing the data for upper layers, transformations were applied to both optical flow and RGB frames, PSL samples were resized to be compliant with ImageNet pre-trained models using a 224 by 224 shape and three channels for color images, data augmentation transformation were not applied due to the nature of the experiment where samples were captured using similar light conditions and camera orientations, PSL samples were transformed to tensors and normalized to floats in the [0, 1] interval, finally the transformed and normalized versions of optical flow and RGB samples were tight together in a tensor structure along with the label for usage at upper layers.
\begin{table}[!htb]
\captionsetup{size=footnotesize}
\begin{tabular}{ p{16em} c}
\toprule
\multicolumn{1}{c}{\textbf{Hyper parameter}} & 
	\multicolumn{1}{c}{\textbf{Value}}\\
\midrule
\textbf{Learning Rate}&	0.001\\
\textbf{No Epochs}&	5\\
\textbf{No Steps per Epoch}&	5\\
\textbf{Batch Size}&	10\\
\textbf{Shuffle Buffer Size}&	50\\
\bottomrule
\end{tabular}
\caption{nSDm training hyper parameters} \label{tab:nSDm-hyperparams}
\end{table}

The \textbf{Wrapper Layer} wraps a CNN network designed for learning visual features from optical flow and a RNN network designed for learning features from series of video frames, the layer receives data from the input pipeline then splits it into optical flow and RGB samples and passes them to the CNN and RNN networks respectively. The logits produced are then concatenated and passed to the dense layer.

The dense layer uses a softmax activation function to convert logits into probabilities used for a correct sign classification. Hyper parameters used during the training process are listed on Table \ref{tab:nSDm-hyperparams}

@@

The \textbf{Opticalflow model} described in Figure \ref{fig:opticalflow-architecture} receives decoded opticalflow samples, it uses a Resnet152 backbone pre-trained with ImageNet, we used a fine tuning transfer learning approach, the backbone produces a 7x7x2048 output that then is passed to a Global Average Pooling layer for obtaining a flattened output of 1x1x2048 which is then passed to a dense layer for logits computation.

@@

The \textbf{Recurrent RGB model} described in Figure \ref{fig:rgb-architecture} receives a sequence of decoded video frames, it uses a \textit{many-to-one} architecture with LSTM cells that hold state of 2048 units length, the output produced by the recurrent layers is then passed to a dense layer for logits computation.

@@

\subsection{Sign detection testing}\label{new-videos-testing}
We selected videos for consultant 24 which will be described in \ref{datasetdesc} for testing nSDm (the model was not trained with consultant 24 videos), we proceed to extract RGB frames for video fragments of five seconds length moving ahead one second on each loop until the end of the video that way we can reduce the number of signs cut or incomplete between video fragments. Optical flow and RGB frames generators feed nSDm for detecting signs \ref{nsdm} present in the video fragment.

@@

\section{Experimentation}\label{experimentation}
\subsection{Dataset Description \cite{lsp_dataset}}\label{datasetdesc}
The PSL dataset was developed by the PUCP Grammar and Signs research group in 2014 and consists in a set of videos recorded during the interviews of 24 individuals, 12 male and 12 female informants, all of them are Lima Peru residents and reported to be born with a permanent deafness condition or acquired the condition before the acquisition of Spanish.
